\chapter*{List of Publications}

\begin{enumerate}
    \item Bogdanov, S., Shepelsky, D., Vasylchenkova, A., \textbf{Sedov, E.}, Freire, P.J., 
    Turitsyn, S.K. and Prilepsky, J.E., 2023. Phase computation for the 
    finite-genus solutions to the focusing nonlinear Schrödinger equation 
    using convolutional neural networks. \textit{Communications in Nonlinear Science and Numerical Simulation}, \textbf{125}, p.107311.
    \item Srivallapanondh, S., Freire, P.J., Alam, A., Costa, N., Spinnler, B., 
    Napoli, A., \textbf{Sedov, E.}, Turitsyn, S.K. and Prilepsky, J.E., 2023. 
    Multi-Task Learning to Enhance Generalizability of Neural Network 
    Equalizers in Coherent Optical Systems. \textit{arXiv preprint arXiv:2307.05374}.
    \item \textbf{Sedov, E.}, 2023, June. Gradient Boosting for Nonlinear Equalization in Optical Transmission Systems. In \textit{2023 Conference on Lasers and Electro-Optics Europe \& European Quantum Electronics Conference (CLEO/Europe-EQEC)} (pp. 1-1). IEEE.
    \item Reznichenko, A.V., Chernykh, A.I., \textbf{Sedov, E.V}. and Terekhov, I.S., 2022.
    Optimal input signal distribution for a nonlinear optical fiber channel
    with small Kerr nonlinearity. \textit{JOSA B}, \textbf{39}(3), pp.810-820.
    \item \textbf{Sedov, E.V.}, Chekhovskoy, I.S. and Prilepsky, J.E.E., 2021. Neural 
    network for calculating direct and inverse nonlinear Fourier transform. \textit{Quantum Electronics}, \textbf{51}(12), p.1118.
    \item \textbf{Sedov, E.V.}, Freire, P.J., Seredin, V.V., Kolbasin, V.A., 
    Kamalian-Kopae, M., Chekhovskoy, I.S., Turitsyn, S.K. and Prilepsky, 
    J.E., 2021. Neural networks for computing and denoising the continuous 
    nonlinear Fourier spectrum in focusing nonlinear Schrödinger equation. \textit{Scientific Reports}, \textbf{11}(1), p.22857.
    \item Chekhovskoy, I.S., Medvedev, S.B., Vaseva, I.A., \textbf{Sedov, E.V.} and 
    Fedoruk, M.P., 2021, June. Fast Eigenvalue Evaluation of the Direct 
    Zakharov-Shabat Problem in Telecommunication Signals Using Adaptive 
    Phase Jump Tracking. In \textit{The European Conference on Lasers and Electro-Optics} (p. ci\_p\_2). Optica Publishing Group.
    \item \textbf{Sedov, E.}, Freire, P.J., Chekhovskoy, I., Turitsyn, S. and Prilepsky, 
    J., 2021, September. Neural Networks For Nonlinear Fourier Spectrum 
    Computation. In \textit{2021 European Conference on Optical Communication (ECOC)} (pp. 1-4). IEEE.
    \item \textbf{Sedov, E.}, Prilepsky, J., Chekhovskoy, I. and Turitsyn, S., 2021, June. 
    Computing continuous nonlinear Fourier spectrum of optical signal with 
    artificial neural networks. In \textit{European Quantum Electronics Conference}. Optica Publishing Group.
    \item Chekhovskoy, I., Medvedev, S.B., Vaseva, I.A., \textbf{Sedov, E.V.} and Fedoruk, 
    M.P., 2021. Introducing phase jump tracking-a fast method for eigenvalue
    evaluation of the direct Zakharov-Shabat problem. \textit{Communications in Nonlinear Science and Numerical Simulation}, \textbf{96}, p.105718.
    \item \textbf{Sedov, E.V.}, Redyuk, A.A., Fedoruk, M.P. and Turitsyn, S.K., 2020, 
    November. Statistical occurrence of soliton content in the conventional 
    optical WDM signals. In \textit{2020 International Conference Laser Optics (ICLO)} (pp. 1-1). IEEE.
    \item \textbf{Sedov, E.V.}, Chekhovskoy, I.S., Prilepsky, J.E.E. and Fedoruk, M.P., 
    2020. Application of neural networks to determine the discrete spectrum 
    of the direct Zakharov–Shabat problem. \textit{Quantum Electronics}, \textbf{50}(12), p.1105.
    \item Turitsyn, S., \textbf{Sedov, E.}, Redyuk, A. and Fedoruk, M., 2019. Nonlinear 
    spectrum of conventional OFDM and WDM return-to-zero signals in 
    nonlinear channel. \textit{Journal of Lightwave Technology}, \textbf{38}(2), pp.352-358.
    \item \textbf{Sedov, E.V.}, Redyuk, A.A., Fedoruk, M.P., Gelash, A.A., Frumin, L.L. and
    Turitsyn, S.K., 2018. Soliton content in the standard optical OFDM 
    signal. \textit{Optics Letters}, \textbf{43}(24), pp.5985-5988.
\end{enumerate}