\chapter*{Acknowledgements}
\markboth{Acknowledgements}{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% put your text here

% My deepest gratitude goes to my supervisor, Professor Sergei Turitsyn, for his unwavering guidance and for making my doctoral journey possible. I extend my heartfelt thanks to my co-supervisor, Dr. Yaroslav Prilepsky, for his assistance and steadfast commitment to research integrity. My colleagues deserve special recognition for revealing to me the genuine essence of academic life. A special thank you to the TRANSNET program and it's PI, Polina Byvel, for supporting my research. I am indebted to my friends, Misha (who passed in 9 days after my thesis defence), Valentine (who first son born in 11 days after my devence), Stepan and many others; to the members of the "BHYK" creativity group (and related): Ivan, Tusnin, Edos, Tuman, Max, Andrey, Tolyan, Alex, Dremov, Shipko, Mishin, Kuprik, Roman, Sanek, Kudashkin, Gervaz, Valera, Rabus for their constant support and for inspiring me to strive for excellence.
% To my family, both present and past, who have nourished my research integrity and nurtured the scientist within me, I owe my deepest thanks.
% Lastly, the most profound thanks are reserved for my wife, Olga, my steadfast companion through life's most rigorous challenges, and for my beloved dog, Bubaleh, who has brought boundless joy and harmony into my life.
My deepest gratitude goes to my supervisor, Professor Sergei Turitsyn, for his unwavering guidance and for making my doctoral journey possible. I extend my heartfelt thanks to my co-supervisor, Dr. Yaroslav Prilepsky, for his assistance and steadfast commitment to research integrity.
My colleagues deserve special recognition for revealing to me the genuine essence of academic life. A special thank you to the TRANSNET program and its Principal Investigator, Polina Bayvel, for supporting my research.

I am indebted to my friends, Misha (who passed away just 9 days after my thesis defense), Valentine (whose first son, Gordei, was born 11 days after my defense), Stepan, and many others; to the members of the "BHYK" creativity group (and related): Ivan, Tusnin, Edos, Tuman, Max, Andrey, Tolyan, Alex, Dremov, Shipko, Mishin, Kuprik, Roman, Sanek, Kudashkin, Gervaz, Valera, Rabus for their constant support and for inspiring me to strive for excellence.

To my parents and family, both present and past, who have nourished my pursuit of knowledge and nurtured the scientist within me, I owe my deepest thanks.

Lastly, the most profound thanks are reserved for my wife, Olga, my steadfast companion through life's most rigorous challenges, and for my beloved dog, Bubaleh, who has brought boundless joy and harmony into my life.


\bigskip
 
\noindent\textit{Birmingham, \today}
\hfill E.~S.
